Run deep learning on AWS instance

### 2021-01-23 docs/ensemble_labels.Rmd
Add acknowledgement

```{r}
library(tidyr)
library(parallel)
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')
tiles <- st_read('data/geoms/tiles_nicfi.geojson',
                 quiet = T) %>% dplyr::select(tile)
tiles <- do.call(rbind, mclapply(1:nrow(tiles), function(n){
  tile <- tiles %>% slice(n)
  tiles_grids <- st_make_grid(tile, n = c(8, 8)) %>% 
    st_sf() %>% arrange() %>% 
    mutate(tile = tile$tile,
                       index = c(1:64))
}, mc.cores = 10))
st_write(tiles, 'data/geoms/tiles_grids.geojson')
tiles <- tiles %>% vect()

tiles_labels <- do.call(bind_rows, mclapply(1:nrow(tiles), function(n){
  labels_each <- crop(lc_labels, tiles[n, ])
  freq(labels_each) %>% data.frame() %>% 
    filter(value != 0) %>%
    mutate(count = count / (dim(labels_each)[1] * dim(labels_each)[2])) %>%
    left_join(., lc_types, by = c('value' = 'id')) %>%
    dplyr::select(-c(value, layer)) %>%
    pivot_wider(names_from = name, values_from = count) %>%
    mutate(tile = tiles[n, ]$tile,
           index = tiles[n, ]$index)
}, mc.cores = 12))

labels_sum <- labels_sum %>% filter(count > 0)
tiles_grids_selected <- do.call(rbind, 
        lapply(labels_sum$name, 
               function(nm){
  if (nm %in% c('Forest', 'Grassland', 'Cropland', 'Shrubland')){
    tile_select <- tiles_labels %>% 
      dplyr::select(tile, index, nm) %>% 
      # arrange_at(nm, funs(desc)) %>% 
      filter(!is.na(!!rlang::sym(nm))) %>% 
      filter((!!rlang::sym(nm)) >= 0.4) %>% 
      mutate(group = ifelse(!!rlang::sym(nm) >= 0.8, 1, 
                            ifelse(!!rlang::sym(nm) >= 0.6, 2, 3))) 
    ns <- tile_select %>% group_by(group) %>% summarise(n = n()) %>% 
      mutate(n = ifelse(n < 1000, n, 1000))
    tile_select <- do.call(rbind, 
                           lapply(unique(tile_select$group), 
                                  function(g){
      set.seed(124 + g)
      tile_select %>% filter(group == g) %>% 
        sample_n(size = ns %>% 
                   filter(group == g) %>% 
                   pull(n))
    }))
  } else if (nm == 'Water'){
    tile_select <- tiles_labels %>% 
      dplyr::select(tile, index, nm) %>% 
      # arrange_at(nm, funs(desc)) %>% 
      filter(!is.na(!!rlang::sym(nm))) %>% 
      filter((!!rlang::sym(nm)) >= 0.4) %>% 
      mutate(group = ifelse(!!rlang::sym(nm) >= 0.8, 1, 
                            ifelse(!!rlang::sym(nm) >= 0.6, 2, 3))) 
    ns <- tile_select %>% group_by(group) %>% summarise(n = n()) %>% 
      mutate(n = ifelse(n < 500, n, 500))
    tile_select <- do.call(rbind, 
                           lapply(unique(tile_select$group), 
                                  function(g){
      set.seed(124 + g)
      tile_select %>% filter(group == g) %>% 
        sample_n(size = ns %>% 
                   filter(group == g) %>% 
                   pull(n))
   }))
  } else {
    # Bare land, wetland, and urban are classes that are hard to tell,
    # and the ensemble labels are less,
    # so we keep as many samples as possible.
    # but also if there are too few labels within a grid,
    # we assume the labels are not reliable anymore.
    # So we use a threshold 0.1.
    tile_select <- tiles_labels %>% 
      dplyr::select(tile, index, nm) %>%
      filter(!is.na(!!rlang::sym(nm))) %>%
      filter((!!rlang::sym(nm)) > 0.1)
  }
  data.frame(name = nm, 
             tile = tile_select$tile, 
             index = tile_select$index)
}))
```

@@

```{r split}
set.seed(1000)
tiles_grids_train <- tiles_grids_selected %>% 
  group_by(name) %>% sample_frac(0.6) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'train') %>%
  dplyr::select(tile, index, usage)
set.seed(1200)
tiles_grids_valid <- tiles_grids_selected %>% 
  anti_join(tiles_grids_train, 
            by = c('tile', 'index')) %>% 
  group_by(name) %>% sample_frac(0.5) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'validate') %>%  
  dplyr::select(tile, index, usage)
tiles_grids_test <- tiles_grids_selected %>%
  dplyr::select(-name) %>% unique() %>% 
  anti_join(tiles_grids_train) %>% 
  anti_join(tiles_grids_valid) %>% 
  mutate(usage = 'test') %>% 
  dplyr::select(tile, index, usage)
catalog_pred_val_test <- rbind(tiles_grids_train, 
                               tiles_grids_valid, 
                               tiles_grids_test)
write.csv(catalog_pred_val_test, 
          'data/intermid/catalog_pred_val_test.csv',
          row.names = T)
```

#### Generate labels images
```{r labels_img}
tiles <- st_read('data/geoms/tiles_grids.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')

@@

lapply(unique(catalog_pred_val_test$tile), function(tile_nm){
  tiles_each <- catalog_pred_val_test %>% filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% aggregate(fact = 512)
  vals <- matrix(1:64, nrow = 8, ncol = 8) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- trim(temp_each) %>% 
      disaggregate(fact = 512)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    writeRaster(label_each, 
                file.path(labels_path, 
                          glue('labels_{tile_nm}_{index}.tif')),
                wopt = list(datatype = 'INT1U',
                        gdal=c("COMPRESS=LZW")))
    rm(tile_each, temp_each, label_each, index);gc()
  })
  rm(tiles_each, lc_labels_each, nm, temp, vals);gc()
})
```

>>>>

```{r}
library(tidyr)
library(parallel)
# Define the width of chips to cut
# In this case, 256 *256 chip.
width_chips <- 16
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')
tiles <- st_read('data/geoms/tiles_nicfi.geojson',
                 quiet = T) %>% dplyr::select(tile)
tiles <- do.call(rbind, mclapply(1:nrow(tiles), function(n){
  tile <- tiles %>% slice(n)
  tiles_grids <- st_make_grid(tile, n = c(width_chips, width_chips)) %>% 
    st_sf() %>% arrange() %>% 
    mutate(tile = tile$tile,
                       index = c(1:(width_chips^2)))
}, mc.cores = 12))
st_write(tiles, 'data/intermid/tiles_grids_256.geojson')
tiles <- tiles %>% vect()

tiles_labels <- mclapply(1:nrow(tiles), function(n){
  labels_each <- crop(lc_labels, tiles[n, ])
  lb <- freq(labels_each) %>% data.frame() %>% 
    filter(value != 0) %>%
    mutate(count = count / (dim(labels_each)[1] * dim(labels_each)[2])) %>%
    left_join(., lc_types, by = c('value' = 'id')) %>%
    dplyr::select(-c(value, layer)) %>%
    pivot_wider(names_from = name, values_from = count) %>%
    mutate(tile = tiles[n, ]$tile,
           index = tiles[n, ]$index)
  if (nrow(lb) > 0) lb
}, mc.cores = 12) %>% bind_rows()
save(tiles_labels, file = 'data/intermid/tiles_labels_256.rda')

labels_sum <- labels_sum %>% filter(count > 0)
tiles_grids_selected <- do.call(rbind, 
                                lapply(labels_sum$name, 
                                       function(nm){
    message(nm)
    if (nm %in% c('Forest', 'Grassland', 'Cropland', 'Shrubland')){
        set.seed(123)
        tile_select <- tiles_labels %>% 
            dplyr::select(tile, index, nm) %>% 
            # arrange_at(nm, funs(desc)) %>% 
            filter(!is.na(!!rlang::sym(nm))) %>% 
            filter((!!rlang::sym(nm)) >= 0.7) %>% 
            sample_n(min(1000, nrow(.)))
    } else if (nm == 'Water'){
        set.seed(566)
        tile_select <- tiles_labels %>% 
            dplyr::select(tile, index, nm) %>% 
            # arrange_at(nm, funs(desc)) %>% 
            filter(!is.na(!!rlang::sym(nm))) %>%
            filter((!!rlang::sym(nm)) <= 0.8) %>%
            filter((!!rlang::sym(nm)) >= 0.4) %>% 
            sample_n(500)
    } else {
        # Bare land, wetland, and urban are classes that are hard to tell,
        # and the ensemble labels are less,
        # so we keep as many samples as possible.
        # but also if there are too few labels within a grid,
        # we assume the labels are not reliable anymore.
        # So we use a threshold 0.1.
        tile_select <- tiles_labels %>% 
            dplyr::select(tile, index, all_of(nm)) %>%
            filter(!is.na(!!rlang::sym(nm))) %>%
            filter((!!rlang::sym(nm)) > 0.1)
    }
    data.frame(name = nm, 
               tile = tile_select$tile, 
               index = tile_select$index)
}))
```

@@

```{r split}
set.seed(1000)
tiles_grids_train <- tiles_grids_selected %>% 
  group_by(name) %>% sample_frac(0.6) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'train') %>%
  dplyr::select(tile, index, usage)
set.seed(1200)
tiles_grids_valid <- tiles_grids_selected %>% 
  anti_join(tiles_grids_train, 
            by = c('tile', 'index')) %>% 
  group_by(name) %>% sample_frac(0.6) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'validate') %>%  
  dplyr::select(tile, index, usage)
tiles_grids_test <- tiles_grids_selected %>%
  dplyr::select(-name) %>% unique() %>% 
  anti_join(tiles_grids_train) %>% 
  anti_join(tiles_grids_valid) %>% 
  mutate(usage = 'test') %>% 
  dplyr::select(tile, index, usage)
catalog_train_val_test <- rbind(tiles_grids_train, 
                               tiles_grids_valid, 
                               tiles_grids_test)
write.csv(catalog_train_val_test, 
          'data/intermid/catalog_train_val_test.csv',
          row.names = F)
```

#### Generate labels images
In order to save the resources for deep learning, we re-defined the class index to be continuous. So:

```
id          name
1       Cropland
2         Forest
3      Grassland
4      Shrubland
5        Wetland
6          Water
7 Urban/Built up (8)
8       Bareland (9)
```

```{r labels_img}
library(glue)
tiles <- st_read('data/intermid/tiles_grids.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')

@@

mclapply(unique(catalog_train_val_test$tile), 
         function(tile_nm){
  tiles_each <- catalog_train_val_test %>% 
    filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% 
    aggregate(fact = 4096/width_chips)
  vals <- matrix(1:(width_chips^2), 
                 nrow = width_chips, 
                 ncol = width_chips) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- terra::trim(temp_each) %>% 
      disaggregate(fact = 4096/width_chips)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    label_each[label_each == 8] <- 7
    label_each[label_each == 9] <- 8
    writeRaster(label_each, 
                file.path(labels_path, 
                          glue('labels_{tile_nm}_{index}.tif')),
                wopt = list(datatype = 'INT1U',
                        gdal=c("COMPRESS=LZW")))
    rm(tile_each, temp_each, label_each, index);gc()
  })
  rm(tiles_each, lc_labels_each, nm, temp, vals);gc()
}, mc.cores = 12)
```

### 2021-01-23 docs/pre_train.Rmd
Add acknowledgement

## PlanetScope quads
plt_quads_rain_season <- list.files('/Volumes/elephant/plt_nicfi',
                        full.names = T, pattern = '2017-12_2018-05')
plt_quads_dry_season <- list.files('/Volumes/elephant/plt_nicfi',
                        full.names = T, pattern = '2018-06_2018-11')
## Labels
lc_labels <- list.files('/Volumes/elephant/labels',
                        full.names = T, pattern = '.tif')

@@

```{r catalog}
library(dplyr)
library(stringr)
library(tidyr)

>>>>

## PlanetScope quads
plt_quads_rain_season <- list.files('/Volumes/elephant/plt_nicfi',
                        full.names = T, pattern = '2017-12_2018-05')
plt_quads_dry_season <- list.files('/Volumes/elephant/plt_nicfi',
                        full.names = T, pattern = '2018-06_2018-11')
## Labels
lc_labels <- list.files('/Volumes/elephant/labels',
                        full.names = T, pattern = '.tif$')

@@

```{r catalog}
library(dplyr)
library(stringr)
library(tidyr)
library(aws.s3)

### 2021-01-23 docs/run_dl_aws.Rmd
Add acknowledgement

Before start, install `jq`.
Then use the scripts in `pytorch_planet/aws_tools/create_spot_instance.sh` using the template:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh <ami_id> <instance_type> <security_group_id> <new_instance_name> <spot_type> <valid_until> <key_name>
```
For example:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh ami-0643df3b347740189 g3.8xlarge sg-0a8bbc91697d6a76b lc_tz persistent 2021-01-20T23:00:00 lsong-keypair
```

@@

## Result visualization

>>>>

Before start, install `jq`.
Then use the scripts in `pytorch_planet/aws_tools/create_spot_instance.sh` using the template:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh <ami_id> <instance_type> <security_group_id> <new_instance_name> <spot_type> <valid_until> <key_name>
```
For example:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh ami-0643df3b347740189 g3.16xlarge sg-0a8bbc91697d6a76b lc_dl_tz persistent 2021-01-22T23:00:00 lsong-keypair
```

@@

## Experiemnts

### Experiment 1

```
Model: Deeplab3+
train_batch = 32
val_batch = 16
transformation = ['vflip', 'hflip', 'rotate', 'resize']
rotate_degree = [-90, 90]
epoch = 30
lr_init = 0.0005
lr_decay = 0.1
lr_decay_step = 1
momentum = 0.99
criterion = BalancedTverskyFocalLoss(ignore_index=0)
optimizer = "adam"
```

**Result**: terminate early. The model performance stuck quickly. Then it got stuck here.

```
train loss: 0.6423281792689253
train acc: 0.7882347084857799
validation loss: 0.872987336890642
validation acc: 0.8824300759060438
time: 409
[14/100]
train loss: 0.6659585610032082
train acc: 0.7867565039131377
validation loss: 0.8747954650153947
validation acc: 0.8814900846675385
time: 391
```

Seems like BalancedTverskyFocalLoss and adam is a good combination of compiler parameters. The problems left is which model and the learning rate. The experiments shown that the complexity of the model is not enough.

### Experiments

```
Model: Unet

train loss: 0.6717923217349582
train acc: 0.7490688176066788
validation loss: 0.8887722446301649
validation acc: 0.8503072788549024
```

```
Model: Deeper Unet

train loss: 0.6798290364720203
train acc: 0.7305447139121868
validation loss: 0.8891407523390858
validation acc: 0.8469047265690427
time: 245
```

```
Model: VGG 
```

```
Model: ResNet 
```

```
Model: DenseNet 
```

```
Model: SegNet 
```
